Constraint Programming - Inside Gecode

### 2014-05-11 paper.tex
corrections

	In constraint programming, search trees, like the one in \reffig{fig:searchTree}, describe the partition of the search space. Their nodes have names, depending on their position in the tree. The search is started from the root node. At this point, all variable domains are still complete. The root node is also the first one with a branch, where the search space is split up into disjoint sets. This approach is similar to the divide-and-conquer technique as it reduces a complex problem into two or more parts with less complexity. The splitting, which adds additional constraints on the variables, is then recursively done with different heuristics to partition the search space into smaller parts. If one of the parts is small enough to decide if it is feasible the node can be called a either a dead end or a solution. To end up as a dead end, there must be no feasible solutions left in the search space at this node. To turn the node into a solution it is necessary that all members of the reduced search space are valid considering the constraints. This can be the case if either only one choice is left for every variable, and the constraints are valid for this choice, or if a whole interval of a floating variable is valid for the constraints.

@@

	Constraints in \emph{Gecode} are evaluated ahead of a further growth of a subtree, as it may be possible to prune some of the sub trees. The problem here is to decide in which order the constraints should be considered, and it may be even necessary to evaluate some constraints more often. A short example will illustrate this \cite[p.~575]{handbookCP}: Let there be two constraints 
	\begin{equation}
		c_1: y - x^2 = 0
	\end{equation}
	\begin{equation}
		c_2: y - x - 1 = 0
		\label{eq:linear}
	\end{equation}
	and let the variable domains be $x, y \in [-4, 4]$. Starting with the evaluation of constraint $c_1$ follows a chain of reductions of the variable domains, based only on constraint propagation:
	\begin{equation}
	\begin{split}
		c_1:\ &y \in [-4, 4] \rightarrow y \in [0, 4] \\
		c_1:\ &x \in [-4, 4] \rightarrow x \in [-2, 2] \\
		c_2:\ &y \in [0, 4] \rightarrow y \in [0, 3] \\
		c_2:\ &x \in [-2, 2] \rightarrow x \in [-1, 2] \\
		c_1:\ &x \in [-1, 2] \rightarrow x \in [-1, 1.73 \dots] \\
		&\dots
	\end{split}
	\end{equation}

@@

	The branching strategy depends on the variable domain: For integer variables there are three common strategies \cite[p.~87]{handbookCP} which differ in how the domain is distributed to the sub trees.
	\begin{itemize}
		\item Enumeration: The domain for $x \in D = {x_1, x_2, \dots}$ is split up into $|D|$-branches where for every branch $i$ one additional constraint $c:\ x = x_i$ is posted.
		\item Binary choice points: One value $x_i$ of the domain $D$ of the variable $x$ is selected and the domain is split up into two sub trees: On one the constraint $x = x_i$ and on the other one the constraint $x \ne x_i$ is posted. This results into something similar to a binary tree at this point.
		\item Domain splitting: The domain $D$ for the variable $x$ is split up into two disjoint intervals or sets, for example $x < 2$ and $x \ge 2$.
	\end{itemize}

	As for variables with a float domain the enumeration and binary choice points are not feasible, the last one, domain splitting, is typically used. As termination criteria for a float it is often necessary to define a certain epsilon $\epsilon > 0$ which describes the interval length, at which the solver states to have found a solution.

@@

	\subsubsection{Depth-First Search}
	The depth-first search was historical the first one which was developed. The idea behind this search algorithm is that every possible combination of variables is examined. In its basic form, it can not be applied to problems with infinite domains; for problems with such variable domains it is necessary to use for instance branch-and-reduce. Sometimes the depth-first search is also referred to as chronological backtrack search.

@@

	For the bounding it is first necessary to initialize a upper bound with a value, for example with $\infty$. This upper bound is updated during the process every time a new optimal solution is found to this new optimum. If a new possible solution is found which is above this upper bound the algorithm can prune it, as only the minimum solution is of interest. The actually pruning of sub problems is based on the relaxed optimum, which is the optimum of the problem without consideration that only discrete values are allowed. If this relaxed optimum for a certain sub problem is greater than the current upper bound it is possible to prune this complete sub problem, as it can not contain a better solution.

@@

	\subsubsection{Limited Discrepancy Search}
	The limited discrepancy search falls into the category of Best-first searches and was first mentioned by Harvey and Ginsberg \cite{limitedDiscrepancy}. It depends on branching heuristics for the tree, which are determined through branching strategies. The basic concept is that during a chronological backtrack search a wrong turn close to the root is first of all very expensive. Second, heuristics work worse the farther away from the solution they are applied, therefore decisions close to the root node are likely to fail. Consequently, in a depth-first search the worst decisions are the most expensive ones. The limited discrepancy search solves exactly this problem as it suggests to change rather the decisions made close to the root than the ones the search made later. The improvement compared to a chronological backtrack was shown by experimental results and theoretically proven in \cite{limitedDiscrepancy}.

	\subsubsection{Backjumping}
	A specific version of backjumping was already mentioned, the limited discrepancy search, where instead of jumping one node back the algorithm always jumps as far back to the root node as possible. In general this is proven to be better than the chronological backtracking \cite{limitedDiscrepancy}, but for certain problems it may be possible to find better nodes to jump back. A jump back in the search tree means to detect a so-called nogood decision, which caused the violation of a constraint in the dead end. How these nogoods must be selected and stored to improve the search depends heavily on the specific field and if it is done wrong the backjumping can even be less efficient than chronological backtracking \cite[p.~100]{handbookCP}.

>>>>

	In constraint programming, search trees, like the one in \reffig{fig:searchTree}, describe the partition of the search space. Their nodes have names, depending on their position in the tree. The search is started from the root node. At this point, all variable domains are still complete. The root node is also the first one with a branch, where the search space is split up into disjoint sets. This approach is similar to the divide-and-conquer technique, as it reduces a complex problem into two or more parts with less complexity. The splitting, which adds additional constraints on the variables, is then recursively done with different heuristics to partition the search space into smaller parts. If one of the parts is small enough to decide if it is feasible the node can be called a either a dead end or a solution. To end up as a dead end, there must be no feasible solutions left in the search space at this node. To turn the node into a solution it is necessary that all members of the reduced search space are valid considering the constraints. This can be the case if either only one choice is left for every variable, and the constraints are valid for this choice, or if a whole interval of a floating variable is valid for the constraints.

@@

	Constraints in \emph{Gecode} are evaluated ahead of a further growth of a subtree, as it may be possible to prune some of the sub trees. The problem here is to decide in which order the constraints should be considered, and it may even be necessary to evaluate some constraints more often. A short example will illustrate this \cite[p.~575]{handbookCP}: Let there be two constraints 
	\begin{equation}
		c_1: y - x^2 = 0
	\end{equation}
	\begin{equation}
		c_2: y - x - 1 = 0
		\label{eq:linear}
	\end{equation}
	and let the variable domains be $x, y \in [-4, 4]$. Starting with the evaluation of constraint $c_1$ follows a chain of reductions of the variable domains, based only on constraint propagation:
	\begin{equation}
	\begin{split}
		c_1:\ &y \in [-4, 4] \rightarrow y \in [0, 4] \\
		c_1:\ &x \in [-4, 4] \rightarrow x \in [-2, 2] \\
		c_2:\ &y \in [0, 4] \rightarrow y \in [0, 3] \\
		c_2:\ &x \in [-2, 2] \rightarrow x \in [-1, 2] \\
		c_1:\ &x \in [-1, 2] \rightarrow x \in [-1, 1.73 \dots] \\
		&\dots
	\end{split}
	\end{equation}

@@

	The branching strategy depends on the variable domain: For integer variables there are three common strategies \cite[p.~87]{handbookCP} which differ in how the domain is distributed to the sub trees.
	\begin{itemize}
		\item Enumeration: The domain for $x \in D = {x_1, x_2, \dots}$ is split up into $|D|$-branches where for every branch $i$ one additional constraint $c:\ x = x_i$ is posted.
		\item Binary choice points: One value $x_i$ from the domain $D$ of the variable $x$ is selected and the domain is split up into two sub trees: On one the constraint $x = x_i$ and on the other one the constraint $x \ne x_i$ is posted. This results into something similar to a binary tree at this point.
		\item Domain splitting: The domain $D$ for the variable $x$ is split up into two disjoint intervals or sets, for example $x < 2$ and $x \ge 2$.
	\end{itemize}

	As for variables with a float domain the enumeration and binary choice points are not feasible, the last one, domain splitting, is typically used. As termination criteria for a float it is necessary to define a certain epsilon $\epsilon > 0$ which describes the interval length, at which the solver states to have found a solution.

@@

	\subsubsection{Depth-First Search}
	The depth-first search was historical the first one which was developed, therefore it is a good point to start explaining search algorithms. The idea behind this search algorithm is that every possible combination of variables is examined. In its basic form, it can not be applied to problems with infinite domains; for problems with such variable domains it is necessary to use for instance branch-and-reduce. Sometimes the depth-first search is also referred to as chronological backtrack search.

@@

	For the bounding it is first necessary to initialize a upper bound with a value, for example with $\infty$. This upper bound is updated during the process every time a new optimal solution is found to this new optimum. If a new possible solution is found which is above this upper bound the algorithm can prune it, as only the minimum solution is of interest.

	The actual pruning of sub problems is based on the relaxed optimum, which is the optimum of the problem without consideration that only discrete values are allowed. If this relaxed optimum for a certain sub problem is greater than the current upper bound it is possible to prune this complete sub problem, as it can not contain a better solution.

@@

	\subsubsection{Limited Discrepancy Search}
	The limited discrepancy search falls into the category of best-first searches and was first mentioned by Harvey and Ginsberg \cite{limitedDiscrepancy}. It depends on branching heuristics for the tree, which are determined through branching strategies. The basic concept is that during a chronological backtrack search a wrong turn close to the root is first of all very expensive. Second, heuristics work worse the farther away from the solution they are applied, therefore decisions close to the root node are likely to fail. Consequently, in a depth-first search the worst decisions are the most expensive ones. The limited discrepancy search solves exactly this problem as it suggests to change rather the decisions made close to the root than the ones the search made later. The improvement compared to a chronological backtrack was shown by experimental results and theoretically proven in \cite{limitedDiscrepancy}.

	\subsubsection{Backjumping}
	A specific version of backjumping was already mentioned, the limited discrepancy search, where instead of jumping one node back the algorithm always jumps as far back to the root node as possible. In general this is proven to be better than the chronological backtracking \cite{limitedDiscrepancy}. However, for certain problems it may be possible to find better nodes to jump back. A jump back in the search tree means to detect a so-called nogood decision, which caused the violation of a constraint in the dead end. How these nogoods must be selected and stored to improve the search depends heavily on the specific field and if it is done wrong the backjumping can even be less efficient than chronological backtracking \cite[p.~100]{handbookCP}.
